FHJ: A Formal Model for Hierarchical Dispatching and Overriding

### 2018-06-04 ECOOP18/sections/Formalization.tex
address marco's suggestions

\subsubsection{Notations}
The metavariables $I, J$ range over interface names; $x$ ranges over variables; $m$ ranges over method names; $e$ ranges over expressions; and $M$ ranges over method declarations. Following Featherweight Java, we assume that the set of variables includes the special variable \kwthis, which cannot be used as the name of an argument to a method. We use the same
conventions as FJ; we write $\overline{I}$ as shorthand for a possibly empty sequence $I_1, ..., I_n$, which may be indexed by $I_i$; and write $\overline{M}$ as shorthand for $M_1 ... M_n$ (with no commas). We also abbreviate operations on pairs of sequences in an obvious way, writing $\overline{I} \; \overline{x}$ for $I_1 \; x_1, ..., I_n \; x_n$, where $n$ is the length of $\overline{I}$ and $\overline{x}$.

\subsubsection{Interfaces}
\yanlin{Bruno: acyclic, clarified herer.}
In order to achieve multiple inheritance, an interface can have a set of 
parent interfaces, where such a set can be empty. Moreover, as usual
in class-based languages, the class hierarchy is acyclic. The interface declaration $\interface{I}{I}{M}$ introduces an interface named $I$ with parent interfaces $\overline{I}$ and a suite of methods $\overline{M}$. The methods of $I$ may either override methods that are already defined in $\overline{I}$ or add new functionality special to $I$, we will illustrate this in more detail later.

\subsubsection{Methods}
Original methods and hierarchically overriding methods share the same syntax in our model for simplicity.
The concrete method declaration $\method{I}{m}{I_x}{x}{J}{e}$ introduces a
method named $m$ with result type $I$, parameters $\overline{x}$ of
type $\overline{I_x}$ and the overriding target $J$. The body of the
method simply includes the returned expression $e$. Notably, we have introduced the
\kwoverride{} keyword for two cases: if the overridden interface is exactly the enclosing
interface itself, then such a method is seen as originally defined; otherwise it is a hierarchical overriding method. 
Note that in an interface $J$, $
I \; m(\overline{I_x} \; \overline{x}) \; {\{} \kwreturn \; e ; {\}} $ is syntactic sugar for $\method{I}{m}{I_x}{x}{J}{e}$, which is the standard way of method definition in Java-like languages. The definition
of abstract methods is written as $\absmethod{I}{m}{I_x}{x}{J}$, which is
similar to a concrete method but without the method body. 
For simplicity, overloading is not modelled for methods, which
implies that we can uniquely identify a method by its name.

@@

\begin{figure*}[t]
\saveSpaceFig
\begin{mathpar}
	\framebox{$ I <: J $} \hspace{.5in} \subid \\
	\subtrans \hspace{.5in} \subextends \\

@@

In \textsc{(T-Invk)}, as the compilation should not be aware
of the dynamic type, it only requires that invoking $m$ is valid for the static type of the
receiver. The result of $\mbody$ contains the interface that provides the most specific implementation,
the parameters and the return type. We use underscore for the return expression, implying that an empty return expression
from an abstract method is acceptable.
% \textsc{(T-PathInvk)} is the typing judgement for a path invocation. Besides the conditions of \textsc{(T-Invk)}, \textsc{(T-PathInvk)} requires the type of receiver to be the subtype of the specified path type. 
% and \textsc{(T-SuperInvk)} checks if the enclosing type directly extends the specified super type.

@@

\subsection{Small-step Semantics and Congruence}
Figure~\ref{fig:smallstep} defines the small-step semantics and
congruence rules of \MIM{}. When evaluating an expression, they
are invoked and produce a single value in
the end. %\haoyuan{we need to be consistent on paragraph upper/lower case.}

\subsubsection{Semantic Rules} \textsc{(S-Invk)} is the only computation rule we need for method invocation.
As a small-step rule and by congruence, it assumes that the receiver and the arguments are already values.
Specifically, the receiver $(J)\new{I}$ indicates the dynamic type $I$
together with the static type $J$. Therefore $\mbody(m, I, J)$ carries out hierarchical dispatching, acquires
the types, the return expression $e_0$ and the interface $I_0$ which provides the most specific method.
Here we use $e_0$ to imply that the return expression is forced to be non-empty because it requires a concrete implementation. Now the
rule reduces method invocation to $e_0$ with substitution.
Parameters are substituted with arguments, and the \lstinline|this| reference is substituted with the receiver,
and in the meanwhile the static types are recorded via annotations. Finally, the return type $I_e$ is put in the front as an annotation.
\subsubsection{Congruence Rules} \textsc{(C-Receiver)}, \textsc{(C-Args)} and \textsc{(C-FReduce)} are natural congruence rules
on receivers, arguments, and cast-expressions, respectively. \textsc{(C-StaticType)} automatically adds an annotation $I$ to the new
object $\new{I}$. \textsc{(C-AnnoReduce)} merges nested upcasts into a single upcast with the outermost type.

>>>>

\subsubsection{Notations}
The metavariables $I, J, K$ range over interface names; $x$ ranges over variables; $m$ ranges over method names; $e$ ranges over expressions; and $M$ ranges over method declarations. Following Featherweight Java, we assume that the set of variables includes the special variable \kwthis, which cannot be used as the name of an argument to a method. We use the same
conventions as FJ; we write $\overline{I}$ as shorthand for a possibly empty sequence $I_1, ..., I_n$, which may be indexed by $I_i$; and write $\overline{M}$ as shorthand for $M_1 ... M_n$ (with no commas). We also abbreviate operations on pairs of sequences in an obvious way, writing $\overline{I} \; \overline{x}$ for $I_1 \; x_1, ..., I_n \; x_n$, where $n$ is the length of $\overline{I}$ and $\overline{x}$.

\subsubsection{Interfaces}
\yanlin{Bruno: acyclic, clarified herer.}
In order to achieve multiple inheritance, an interface can have a set of 
parent interfaces, where such a set can be empty. Moreover, as usual
in class-based languages, \yanlin{refined 'acyclic'} the extension relation over interfaces is acyclic. The interface declaration $\interface{I}{I}{M}$ introduces an interface named $I$ with parent interfaces $\overline{I}$ and a suite of methods $\overline{M}$. The methods of $I$ may either override methods that are already defined in $\overline{I}$ or add new functionality special to $I$, we will illustrate this in more detail later.

\subsubsection{Methods}
Original methods and hierarchically overriding methods share the same syntax in our model for simplicity.
The concrete method declaration $\method{I}{m}{I_x}{x}{J}{e}$ introduces a
method named $m$ with result type $I$, parameters $\overline{x}$ of
type $\overline{I_x}$ and the overriding target $J$. The body of the
method simply includes the returned expression $e$. Notably, we have introduced the
\kwoverride{} keyword for two cases: if the overridden interface is exactly the enclosing
interface itself, then such a method is seen as originally defined \yanlin{added an explanation} (including the case of merging methods from different branches,  which can be viewed as \emph{reborn}); otherwise it is a hierarchical overriding method. 
Note that in an interface $J$, $
I \; m(\overline{I_x} \; \overline{x}) \; {\{} \kwreturn \; e ; {\}} $ is syntactic sugar for $\method{I}{m}{I_x}{x}{J}{e}$, which is the standard way of method definition in Java-like languages. The definition
of abstract methods is written as $\absmethod{I}{m}{I_x}{x}{J}$, which is
similar to a concrete method but without the method body. 
For simplicity, overloading is not modelled for methods, which
implies that we can uniquely identify a method by its name.

@@

\begin{figure*}[t]
\saveSpaceFig
\begin{mathpar}
	\framebox{$ I <: J $} \hspace{.5in} \subid \\
	\subtrans \hspace{.5in} \subextendsForall \\

@@

In \textsc{(T-Invk)}, as the compilation should not be aware
of the dynamic type, it only requires that invoking $m$ is valid for the static type of the
receiver. The result of $\mbody$ contains the interface that provides the most specific implementation,
the parameters and the return type. We use underscore for the return expression, matching both implemented and abstract methods. \yanlin{refined.}
% \textsc{(T-PathInvk)} is the typing judgement for a path invocation. Besides the conditions of \textsc{(T-Invk)}, \textsc{(T-PathInvk)} requires the type of receiver to be the subtype of the specified path type. 
% and \textsc{(T-SuperInvk)} checks if the enclosing type directly extends the specified super type.

@@

\subsection{Small-step Semantics and \Congruence{}}
Figure~\ref{fig:smallstep} defines the small-step semantics and
\congruence{} rules of \MIM{}. When evaluating an expression, they
are invoked and produce a single value in
the end. %\haoyuan{we need to be consistent on paragraph upper/lower case.}

\subsubsection{Semantic Rules} \textsc{(S-Invk)} is the only computation rule we need for method invocation.
As a small-step rule and by congruence, it assumes that the receiver and the arguments are already values.
Specifically, the receiver $(J)\new{I}$ indicates the dynamic type $I$
together with the static type $J$. Therefore $\mbody(m, I, J)$ carries out hierarchical dispatching, acquires
the types, the return expression $e_0$ and the interface $I_0$ which provides the most specific method.
Here we use $e_0$ to imply that the return expression is forced to be non-empty because it requires a concrete implementation. Now the
rule reduces method invocation to $e_0$ with substitution.
Parameters are substituted with arguments, and the \lstinline|this| reference is substituted with the receiver,
and in the meanwhile the static types are recorded via annotations. Finally, the return type $I_e$ is put in the front as an annotation.
\subsubsection{\Congruence{} Rules} \textsc{(C-Receiver)}, \textsc{(C-Args)} and \textsc{(C-FReduce)} are natural \congruence{} rules
on receivers, arguments, and cast-expressions, respectively. \textsc{(C-StaticType)} automatically adds an annotation $I$ to the new
object $\new{I}$. \textsc{(C-AnnoReduce)} merges nested upcasts into a single upcast with the outermost type.

### 2018-06-04 ECOOP18/sections/Overview.tex
address marco's suggestions

\begin{lstlisting}
interface Deck { void draw() {...} }
interface Drawable { void draw() {...} }
interface DrawableDeck extends Drawable, Deck {}
// main program
((Deck) new DrawableDeck()).draw()  // calls Deck.draw
// (new DrawableDeck()).draw()        // this call is ambiguous and rejected
\end{lstlisting}
In our language, the main program is merely an expression. The above cast indicates
that we expect to invoke the \lstinline|draw| method from the branch
\lstinline|Deck|. Similarly, we could have used an upcast to \lstinline|Drawable|
to call the \lstinline|draw| method from \lstinline|Drawable|.
Without the cast, the call would be ambiguous and
\MIM{}'s type system would reject it.

>>>>

\begin{lstlisting}
interface Deck { void draw() {...} }
interface Drawable { void draw() {...} }
interface DrawableDeck extends Drawable, Deck {}
// main program
((Deck) new DrawableDeck()).draw()  // calls Deck.draw
// new DrawableDeck().draw()        // this call is ambiguous and rejected
\end{lstlisting}
\yanlin{Refined.}In our language, a program consists of interfaces declarations and a main an expression which produces the final result. In the above main expression \lstinline|((Deck) new DrawableDeck()).draw()|, the cast indicates
that we expect to invoke the \lstinline|draw| method from the branch
\lstinline|Deck|. Similarly, we could have used an upcast to \lstinline|Drawable|
to call the \lstinline|draw| method from \lstinline|Drawable|.
Without the cast, the call would be ambiguous and
\MIM{}'s type system would reject it.
